---
title: "Reproducible Research: Peer Assessment 1 - Peter Suhner"
output: 
  html_document: 
    keep_md: true
---

## Loading and preprocessing the data
1._Load the data (i.e. read.csv())
   (the data is expected to reside in current directory under the name of "activity.csv")
```{r}
require("knitr")
require("timeDate")

Wrk_Path <- "S:/D/_Data_Management_Strategical/Data_Science/_COURSERA_STUFF/Course_5_Reproducible_Research/Peer_Assessment_1"
In_Path <- paste(Wrk_Path, "InputData", sep="/")
In_File <- paste(In_Path, "activity.csv", sep="/")
act <- read.csv(In_File)
```
- A quick overview of the data structure and contents:
  - a summary of the data frame
  - first and last 5 rows of data content
  - number of complete cases
```{r}
summary(act)
rbind(head(act, 5), tail(act, 5))
sum(complete.cases(act))
```

2._Process/transform the data (if necessary) into a format suitable for the analysis
- aggregation of data:
  - steps per day
  - steps per 5 minutes interval
```{r}
steps_d      <- aggregate(steps ~ date, data=act, FUN=sum)
head(steps_d, 5)
steps_i <- aggregate(steps ~ interval, data=act, FUN=mean)
head(steps_i, 5)
```

## What is mean total number of steps taken per day?
For this part of the work, missing values in the dataset are ignored.

1._Histogram of the <b>total number of steps</b> 
   taken <b>each day</b>

```{r}
barplot(steps_d$steps, names.arg=steps_d$date, xlab="Date", ylab="# Steps")
```

2._Calculate and report the <b>mean</b> 
   and <b>median total number of steps</b> 
   taken <b>per day</b>
```{r}
mean(steps_d$steps)
median(steps_d$steps)
```

## What is the average daily activity pattern?
1._Time series plot (i.e. type = "l") of the <b>5-minute interval</b> (x-axis) 
   and the <b>average number of steps taken</b>, 
   <b>averaged across all days</b> (y-axis)
```{r}
plot(steps_i, type="l")
```


2._Which <b>5-minute interval</b>, 
   on average across all the days in the dataset, 
   contains the <b>maximum number of steps</b>?
   - fetch the element number and display the contents 
```{r}
max_steps_element <- which.max(steps_i$steps)
steps_i[max_steps_element, ]
```


## Imputing missing values
-  Calculate and report the total <b>number of missing values</b> in the
   dataset (i.e. the total number of rows with `NA`s)

```{r}
sum(!complete.cases(act))
```
-  Devise a <b>strategy for filling in all of the missing values</b> in the
   dataset. The strategy does not need to be sophisticated. For
   example, you could use the mean/median for that day, or the mean
   for that 5-minute interval, etc.

     - <b><i>The solution will use the mean for the 5-minute interval as replacement 
             for missing values.</i></b>

-  Create a new dataset that is equal to the original dataset but with
   the missing data filled in.

```{r}
## merge original activity data frame with interval data frame
my_act <- merge(act, steps_i, by = 'interval', all.y = FALSE)

## merge NA values with averages rounding up for integers
my_act$steps.x[is.na(my_act$steps.x)] <- as.integer(round(my_act$steps.y[is.na(my_act$steps.x)]))

## drop obsolete column by renaming to original names and reordering identical to original data frame
names(my_act) <- c("interval", "steps","date","obsolete")
my_act <- my_act[names(act)]
```

-  Make a histogram of the total number of steps taken each day 
   and calculate and report the mean and median total number of steps taken per day.
   
```{r}
my_steps_d      <- aggregate(steps ~ date, data=my_act, FUN=sum)
barplot(my_steps_d$steps, names.arg=my_steps_d$date, xlab="Date", ylab="# Steps")
mean(my_steps_d$steps)
median(my_steps_d$steps)
```

- Do these values differ from the estimates from the first part of the assignment? 
  - <i>I cannot see a major difference between the two patterns</i>
  - <i>However, the graph now shows <b>additional days</b> which were omitted in the original data
    due to the fact that they only contained N/A's</i>

- What is the impact of imputing missing data on the estimates of the total daily number of steps?
  - <i>The result tends to be more complete and aligned as entire days of N/A data are now replaced
    by extrapolated data</i>
  - <i>With no dramatic change on neither the historgram nor the mean/median values of the such
    completed data, we can assume that the overall meaning of the data is not falsified by 
    the approach taken</i>


## Are there differences in activity patterns between weekdays and weekends?
- Solution makes use of "timeDate" R package (containing isWeekend/isWeekday functions)

```{r}
my_act$wknd <- isWeekend(my_act$date)
my_act$wknd <- gsub("FALSE", "Weekday", my_act$wknd)
my_act$wknd <- gsub("TRUE" , "Weekend", my_act$wknd)


par(col="blue", mfrow=c(2,1))
daytypes <- c("Weekend", "Weekday")
for (i in daytypes) {
    steps.i <- aggregate(steps ~ interval,
                           data=my_act,
                           subset=my_act$wknd==i,
                           FUN=mean)
    plot(steps.i, type="l", main=i)
}


```
- <i>Conclusion: We can see a clear difference between weekend and weekday activities</i>
