title: "Reproducible Research: Peer Assessment 1 - Peter Suhner"
output: 
  html_document: 
    keep_md: true
---

## Loading and preprocessing the data
1. Load the data (i.e. read.csv())
   (the data is expected to reside in current directory under the name of "activity.csv")
```{r}
require("knitr")

Wrk_Path <- "S:/D/_Data_Management_Strategical/Data_Science/_COURSERA_STUFF/Course_5_Reproducible_Research/Peer_Assessment_1"
In_Path <- paste(Wrk_Path, "InputData", sep="/")
In_File <- paste(In_Path, "activity.csv", sep="/")
act <- read.csv(In_File)
```
- A quick overview of the data structure and contents
```{r}
summary(act)
rbind(head(act, 5), tail(act, 5))
sum(complete.cases(act))
```

2.Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
steps_d      <- aggregate(steps ~ date, data=act, FUN=sum)
head(steps_d, 5)
steps_i <- aggregate(steps ~ interval, data=act, FUN=mean)
head(steps_i, 5)
```

## What is mean total number of steps taken per day?
For this part of the work, missing values in the dataset are ignored.

1. Histogram of the <b>total number of steps</b> 
   taken <b>each day</b>

```{r}
barplot(steps_d$steps, names.arg=steps_d$date, xlab="Date", ylab="# Steps")
```

2. Calculate and report the <b>mean</b> 
   and <b>median total number of steps</b> 
   taken <b>per day</b>
```{r}
mean(steps_d$steps)
median(steps_d$steps)
```

## What is the average daily activity pattern?
1. Time series plot (i.e. type = "l") of the <b>5-minute interval</b> (x-axis) 
   and the <b>average number of steps taken</b>, 
   <b>averaged across all days</b> (y-axis)
```{r}
plot(steps_i, type="l")
```


2. Which <b>5-minute interval</b>, 
   on average across all the days in the dataset, 
   contains the <b>maximum number of steps</b>?
   - fetch the element number and display the contents 
```{r}
max_steps_element <- which.max(steps_i$steps)
steps_i[max_steps_element, ]
```


## Imputing missing values
1. Calculate and report the total <b>number of missing values</b> in the
   dataset (i.e. the total number of rows with `NA`s)

```{r}
sum(!complete.cases(act))
```
2. Devise a <b>strategy for filling in all of the missing values</b> in the
   dataset. The strategy does not need to be sophisticated. For
   example, you could use the mean/median for that day, or the mean
   for that 5-minute interval, etc.

